---
title: "Hands-on Exercise 01: Geospatial Data Wrangling with R"
execute:
  warning: False
date: "`r Sys.Date()`"
---

# Getting started

## Installing and Loading R Packages

In this section, I will install and load tidyverse and sf packages.

```{r}
pacman::p_load(tidyverse, sf)
```

# Importing a geospatial data in shapefile format

```{r}

mpsz <- st_read(dsn = "data/geospatial",
                layer = "MP14_SUBZONE_WEB_PL")

```

# Importing polyline feature data in shapefile form

```{r}
cyclingpath = st_read(dsn="data/geospatial",
                      layer= "CyclingPathGazette") 

```

# Importing GIS data in kml format

```{r}
preschool=st_read("data/geospatial/PreSchoolsLocation.kml")

```

# Working with functions
## st_geometry()

```{r}
st_geometry(mpsz)

```

## glimpse()
```{r}
glimpse(mpsz)

```

## head()
```{r}
head(mpsz, n=5)

```

# Plotting the Geospatial Data
## Basic Plotting 

```{r}
plot(mpsz)

```

## Plotting st_geometry()
```{r}
plot(st_geometry(mpsz))

```

## Plotting specific attribute

```{r}
plot(mpsz["PLN_AREA_N"])

```


# Projection for Geospatial

## Assigning EPSG code to a simple feature dataframe
###Seeing the CRS
```{r}
st_crs(mpsz)

```

### Setting the CRS to 3414
```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

## Check the CRS again
```{r}
st_crs(mpsz3414)

```

# Transforming the projection of (geospatial) preschool from wgs84 to svy21
## Check the geometry
Firstly, I will check the projection for preschool

```{r}
st_geometry(preschool)
```

## Transform
Then, I will transform the CRS

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)

```

## Display the new CRS

```{r}
st_geometry(preschool3414)

```


# Importing and Converting Aspatial Data
## Reading the csv file
```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

## Reading the listings

```{r}
list(listings)

```


## Convert the simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

## Examine contents
```{r}
glimpse(listings_sf)
```


# Geoprocessing with sf package
## Buffering
### Computing the 5-metres buffers around cycling paths

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

### Calculating area
```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)

```

### Derive total land involved
```{r}
sum(buffer_cycling$AREA)
      
```

# Point in polygon Count
## Calculate number of preschools that fall insde each planning subzone
```{r}
mpsz3414$'PreSch Count'<- lengths(st_intersects(mpsz3414, preschool3414))
```

## Summary of the Count
```{r}
summary(mpsz3414$'PreSch Count')
```

## Subzone with the highest number of pre-schools (ak prof about wt problem)
```{r}
top_n(mpsz3414, 1)

```

# DIY Calculate Density of Pre-school by Planning Subzone
## Derive area of each planning subzone
```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()

```

## Compute density









