---
title: "Hands-on Exercise 12: Modelling Geographical Accessibility"
format:
  html:
    code-fold: True
    code-summary: "Show the code"
    toc: True
    toc-depth: 4
execute:
  eval: True
  echo: True
  warning: False
date: "`r Sys.Date()`"
---

# The data

Four data sets will be used in this hands-on exercise, they are:

-   `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

-   `hexagons`: A 250m radius hexagons GIS data. This data set was created by using [*st_make_grid()*](https://r-spatial.github.io/sf/reference/st_make_grid.html) of sf package. It is in ESRI shapefile format.

-   `ELDERCARE`: GIS data showing location of eldercare service. [This data](https://data.gov.sg/dataset/eldercare-services) is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

-   `OD_Matrix`: a distance matrix in csv format. There are six fields in the data file. They are:

    -   `origin_id`: the unique id values of the origin (i.e. `fid` of hexagon data set.),

    -   `destination_id`: the unique id values of the destination (i.e. `fid` of `ELDERCARE` data set.),

    -   `entry_cost`: the perpendicular distance between the origins and the nearest road),

    -   `network_cost`: the actual network distance from the origin and destination,

    -   `exit_cost`: the perpendicular distance between the destination and the nearest road), and

    -   `total_cost`: the summation of `entry_cost`, `network_cost` and `exit_cost`.

All the values of the cost related fields are in **metres**.

# Loading the Packages

The code chunk below installs and launches these R packages into RStudio environment.

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

# Geospatial Data Wrangling

## Importing geospatial data

```{r}
# Importing the 2014 subzone no sea
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")

# Importing hexagons
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 

# Importing eldercare
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 

```

## Updating CRS Information

The code chunk below updates the newly imported mpsz with the correct ESPG code (i.e. 3414)

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

I can now check the crs of **mpsz** using the `st_crs()` function of the *sf* package.

```{r}
st_crs(mpsz)
```

# Cleaning and updating attribute fields of the geospatial data

There are many redundant fields in the data tables of both `eldercare` and `hexagons`. The code chunks below will be used to exclude those redundant fields. At the same time, a new field called `demand` and a new field called `capacity` will be added into the data table of `hexagons` and `eldercare` sf data frame respectively. Both fields are derived using `mutate()` of *dplyr* package.

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)

hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

# Aspatial Data Handling and Wrangling

## Importing Distance Matrix

The code chunk below uses `read_cvs()` of *readr* package to import `OD_Matrix.csv` into RStudio. The imported object is a tibble data.frame called `ODMatrix`.

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

## Tidying Distance Matrix

The code chunk below uses `spread()` of tidyr package is used to transform the O-D matrix from a thin format into a fat format.

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

::: callout-note
You may now use the pivot_wider() function from the tidyr package
:::

Currently, the distance is measured in metre because SVY21 projected coordinate system is used. The code chunk below will be used to convert the unit f measurement from metre to kilometre.

```{r}
distmat_km <- as.matrix(distmat/1000)
```

# Modelling and Visualising Accessibility using Hansen Method

## Computing Hansen’s accessibility

The code chunk below calculates Hansen’s accessibility using `ac()` of *SpatialAcc* and `data.frame()` is used to save the output in a data frame called **acc_Handsen**.

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

The default field name is very messy, we will rename it to accHansen by using the code chunk below.

```{r}
colnames(acc_Hansen) <- "accHansen"
```

Next, we will convert the data table into tibble format by using the code chunk below.

```{r}
acc_Hansen <- as_tibble(acc_Hansen)
```

Lastly, `bind_cols()` of *dplyr* will be used to join the acc_Hansen tibble data frame with the hexagons simple feature data frame. The output is called **hexagon_Hansen**.

```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

Actually, the steps above can be perform by using a single code chunk as shown below.

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 0.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## Visualising Hansen’s accessibility

### Extracting map extent

Firstly, we will extract the extend of `hexagons` simple feature data frame by by using `st_bbox()` of *sf* package.

```{r}
mapex <- st_bbox(hexagons)
```

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation

In this section, we are going to compare the distribution of Hansen’s accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into **haxegon_Hansen** simple feature data frame by using the code chunk below.

```{r}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

Next, `ggplot(`) will be used to plot the distribution by using boxplot graphical method.

```{r}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# Modelling and Visualising Accessibility using KD2SFCA Method

## Computing KD2SFCA’s accessibility

In this section, you are going to repeat most of the steps you had learned in previous section to perform the analysis. However, some of the codes will be combined into one code chunk. The code chunk below calculates Hansen’s accessibility using `ac()` of SpatialAcc and `data.frame()` is used to save the output in a data frame called acc_KD2SFCA. Notice that KD2SFCA is used for family argument.

```{r}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))

colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- as_tibble(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## Visualising KD2SFCA’s accessibility

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore. Notice that mapex is reused for bbox argument.

```{r}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation

Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into hexagon_KD2SFCA simple feature data frame by using the code chunk below.

```{r}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

Next, `ggplot()` will be used to plot the distribution by using boxplot graphical method.

```{r}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method

## Computing SAM accessibility

In this section, you are going to repeat most of the steps you had learned in previous section to perform the analysis. However, some of the codes will be combined into one code chunk.

The code chunk below calculates Hansen’s accessibility using ac() of SpatialAcc and data.frame() is used to save the output in a data frame called acc_SAM. Notice that SAM is used for family argument.

```{r}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- as_tibble(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## Visualising SAM’s accessibility

The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore. Notice that mapex is reused for bbox argument.

```{r}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical graphic visualisation

Now, we are going to compare the distribution of SAM accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into hexagon_SAM simple feature data frame by using the code chunk below.

```{r}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

Next, `ggplot()` will be used to plot the distribution by using boxplot graphical method.

```{r}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```
